Modal Abstractions for Virtualizing Memory Addresses

### 2023-04-13 programlogic.tex
+

\begin{figure*}
  \begin{lstlisting}[language=Coq]
 Definition $\pfpointsto\locsf\vpts\qfrac\ppts$ :vProp $\Sigma$ := pmapsto_def_v (pfn w) (page_off w) q v.
 Definition $\vaddr\mapsto_{\textsf{t}}\{\textsf{q}\}\; \vpage$ : vProp $\Sigma$ := 
  $\exists_{\textsf{l4e l3e l2e l1}} \ldotp$  $\ulcorner$ aligned $\vaddr \urcorner \ast$ L4_L1_PointsTo(rtv maddr l4e l3e l2e l1e) $\ast$ addr_L1$(\vaddr,\textsf{l1e}) \mapsto_{p} \vpage$.
 Definition L4_L1_PointsTo (rtv maddr l4e l3e l2e l1e :word 64) : vProp $\Sigma$ :=
  (l4M52 maddr rtv) \$\sim$\ (l4off maddr rtv) $\mapsto_{a}$ {q1}  l4e  $\ast$
  (l3M52 maddr l4e) \$\sim$\ (l3off maddr l4e)  $\mapsto_{a}$ {q2}  l3e $\ast$ 
  (l2M52 maddr l3e) \$\sim$\ (l2off maddr l3e) $\mapsto_{a}$ {q3}  l2e $\ast$
  (l1M52 maddr l2e) \$\sim$\ (l1off maddr l2e) $\mapsto_{a}$ {q4}  l1e.
\end{lstlisting}
\caption{A Strong Virtual Points-to Relation}
\todo[inline,color=red]{Colin says: shouldn't ther be a fraction q in the body of the virtual pointsto def?
I know this is just a sketch that we then move away from, but it's worth getting accurate.}
\todo[inline,color=cyan]{Colin says: the virtual points-to here is not defined as a vprop! the body is an iProp,
containing an unbound variable rtv, which should be the cr3val taken as an argument. (Also, rtv doesn't make sense to me
as a shorthand, why not cr3val?). I'm also pretty sure the maddr in the virtual points-to definition should be va.}
  \label{fig:strongvirtualpointsto}
\end{figure*}

@@

\begin{figure*}
\[
\begin{array}{l}
    \vaddr\mapsto_{\textsf{v,rtv}}\{\textsf{q}\}\vpage \stackrel{\triangle}{=} 
    \exists \paddr\ldotp
    \sumapaces\rtv\delta \ast 
  \sumwalkabs\vaddr\qfrac\paddr \ast \paddr \mapsto_{p} \vpage
\end{array}
\]
\todo[inline]{Colin says: again, this is supposed to be a vprop, but isn't doing anything with a cr3val argument.
I realize the GASI and per-address-space invariants, and therefore machinery to tie the virtual points-to
to the GASI, hasn't been introduced yet. Is there a short-hand here we can use to be ``more complete'' until the later text
that fills in these details?
}
\todo[inline,color=cyan]{Colin says: Oh!!! I see. You have that extremely subtle rtv subscript on the virtual points-to
syntax! Don't do that. First, it's really subtle and wasn't explained anywhere, so even I missed it. Second,
we're very explicitly describing vProps as functions from root to iProp. Sticking the rtv on the arrow there
is partly weird because it obscures which argument is really the vProp argument (there are 2 other word64s in that syntax),
partly weird because it's written next to something that \emph{isn't} a variable (the v subscript on the virtual points-to
arrow) while using the same font, and partly weird because when you actually write these assertions, rtv doesn't show
up in the syntax at all, which is the whole point --- virtual points-to defs shouldn't mention their cr3 or address space at
all!
Here's what I think it should look like (I also existentially quantified $\delta$ and fixed the p subscript on physical
points-to):
}
\[
\begin{array}{l}
    \vaddr\mapsto_{\textsf{v}}\{\textsf{q}\}\vpage : \mathsf{vProp}~\Sigma \stackrel{\triangle}{=} 
    \lambda \mathit{cr3val}\ldotp
    \exists \paddr\ldotp
    \exists \delta\ldotp
    \sumapaces{\mathit{cr3val}}\delta \ast 
  \sumwalkabs\vaddr\qfrac\paddr \ast \paddr \mapsto_{\mathsf{p}} \vpage
\end{array}
\]
\todo[inline]{Colin says: The fact that we can't visually distinguish the variable v that is the value in memory
from the subscript v on the virtual points-to arrow is not ideal: it makes it hard to tell what's concrete syntax
and what's a variable. This is worth fixing, but near the bottom of the priority list}
\caption{Virtual-Pointsto for Sharing Pages}
  \label{fig:virtualpointstosharing}
\end{figure*}

@@

  \begin{figure*}
\[
\begin{array}{l}
  \mathcal{I}\textsf{ASpace} \stackrel{\triangle}{:=} 
  \exists\;\gammaPred \;\ldotp\; \ownGhost\gammaPred{\authfull{\ptableabswalk\ptablestore}} \ast \\
  \bigast{(\vaddr, \textsf{l1e})\in \ptablestore}{\exists\;(\textsf{l4e l3e l2e})\ldotp \textsf{L}_{4}\_\textsf{L}_{1}\_\textsf{PointsTo}(\rtv,\textsf{ maddr, l4e, l3e, l2e, l1e})}
\end{array}
\]
\caption{Global Address-Space Invariant}
  \label{fig:peraspaceinvariant}
  \end{figure*}

>>>>

\begin{figure*}
  \begin{lstlisting}[language=Coq]
 Definition $\vaddr\mapsto_{\textsf{t}}\{\textsf{q}\}\; \vpage$ : vProp $\Sigma$ := 
  $\exists_{\textsf{l4e l3e l2e l1}} \ldotp$  $\ulcorner$ aligned $\vaddr \urcorner \ast$ L4_L1_PointsTo(maddr l4e l3e l2e l1e) $\ast$ addr_L1$(\vaddr,\textsf{l1e}) \mapsto_{p} \vpage$.
 Definition L4_L1_PointsTo (rtv maddr l4e l3e l2e l1e :word 64) : vProp $\Sigma$ :=
  (l4M52 maddr rtv) \$\sim$\ (l4off maddr rtv) $\mapsto_{a}$ {q1}  l4e  $\ast$
  (l3M52 maddr l4e) \$\sim$\ (l3off maddr l4e)  $\mapsto_{a}$ {q2}  l3e $\ast$ 
  (l2M52 maddr l3e) \$\sim$\ (l2off maddr l3e) $\mapsto_{a}$ {q3}  l2e $\ast$
  (l1M52 maddr l2e) \$\sim$\ (l1off maddr l2e) $\mapsto_{a}$ {q4}  l1e.
\end{lstlisting}
\caption{A Strong Virtual Points-to Relation}
\todo[inline,color=red]{Colin says: shouldn't ther be a fraction q in the body of the virtual pointsto def?
I know this is just a sketch that we then move away from, but it's worth getting accurate.}
\todo[inline,color=cyan]{Colin says: the virtual points-to here is not defined as a vprop! the body is an iProp,
containing an unbound variable rtv, which should be the cr3val taken as an argument. (Also, rtv doesn't make sense to me
as a shorthand, why not cr3val?). I'm also pretty sure the maddr in the virtual points-to definition should be va.}
  \label{fig:strongvirtualpointsto}
\end{figure*}

@@

\begin{figure*}
\todo[inline]{Colin says: again, this is supposed to be a vprop, but isn't doing anything with a cr3val argument.
I realize the GASI and per-address-space invariants, and therefore machinery to tie the virtual points-to
to the GASI, hasn't been introduced yet. Is there a short-hand here we can use to be ``more complete'' until the later text
that fills in these details?
}
\todo[inline,color=cyan]{Colin says: Oh!!! I see. You have that extremely subtle rtv subscript on the virtual points-to
syntax! Don't do that. First, it's really subtle and wasn't explained anywhere, so even I missed it. Second,
we're very explicitly describing vProps as functions from root to iProp. Sticking the rtv on the arrow there
is partly weird because it obscures which argument is really the vProp argument (there are 2 other word64s in that syntax),
partly weird because it's written next to something that \emph{isn't} a variable (the v subscript on the virtual points-to
arrow) while using the same font, and partly weird because when you actually write these assertions, rtv doesn't show
up in the syntax at all, which is the whole point --- virtual points-to defs shouldn't mention their cr3 or address space at
all!
Here's what I think it should look like (I also existentially quantified $\delta$ and fixed the p subscript on physical
points-to):
}
\[
\begin{array}{l}
    \vaddr\mapsto_{\textsf{v}}\{\textsf{q}\}\vpage : \mathsf{vProp}~\Sigma \stackrel{\triangle}{=} 
    \lambda \mathit{cr3val}\ldotp
    \exists \paddr\ldotp
    \exists \delta\ldotp
    \sumapaces{\mathit{cr3val}}\delta \ast 
  \sumwalkabs\vaddr\qfrac\paddr \ast \paddr \mapsto_{\mathsf{p}} \vpage
\end{array}
\]
\todo[inline]{Colin says: The fact that we can't visually distinguish the variable v that is the value in memory
from the subscript v on the virtual points-to arrow is not ideal: it makes it hard to tell what's concrete syntax
and what's a variable. This is worth fixing, but near the bottom of the priority list}
\caption{Virtual-Pointsto for Sharing Pages}
  \label{fig:virtualpointstosharing}
\end{figure*}

@@

  \begin{figure*}
\[
\begin{array}{l}
  \mathcal{I}\textsf{ASpace}(\ptablestore,m)\stackrel{\triangle}{:=} \exists\;\gammaPred \ldotp \; \lambda\textsf{ rtv} \ldotp \; \ulcorner m \; !!\; \textsf{rtv} = \textsf{Some } \gammPred \urcorner \ast
    \ownGhost\gammaPred{\authfull{\ptableabswalk\ptablestore}} \ast \\
  \bigast{(\vaddr, \textsf{l1e})\in \ptablestore}{\exists\;(\textsf{l4e l3e l2e})\ldotp \textsf{L}_{4}\_\textsf{L}_{1}\_\textsf{PointsTo}(\textsf{ maddr, l4e, l3e, l2e, l1e})}
\end{array}
\]
\caption{Global Address-Space Invariant with a fixed global map of address-space names $m$}
  \label{fig:peraspaceinvariant}
  \end{figure*}
